---
title: "Filter intervals below threshold"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE}

rm(list=ls())

library(tidyverse)

```

The default atacseq pipeline threshold includes very small peaks. The option --macs_fdr 0.00001 did not filter as expected so subsequently filter peaks as follows.

### Set dirs
```{r}

workingdir="/Users/delasj/Documents/BriscoeLab/project_DV_ATAC_reproduce_analysis/"
subworkinput="inputs_cats-atac_1/"
outdir="outputs_cats-atac_1/"

ifelse(!dir.exists(file.path(workingdir,outdir)), dir.create(file.path(workingdir,outdir)), "Directory exists")

```


## Load data

The boolean file from nf-core/atacseq `cats_atac/results/bwa/mergedLibrary/macs/broadPeak/consensus/consensus_peaks.mLb.clN.boolean.annotatePeaks.txt`
  

```{r }

peaks_boolean <- read.delim(file=paste0(workingdir,subworkinput,"consensus_peaks.mLb.clN.boolean.annotatePeaks.txt"),header=TRUE, stringsAsFactors = FALSE)

```

Clean empty columns and column names

```{r}

peaks_boolean_clean <- peaks_boolean %>%
  select(chr,start,end,interval_id,num_peaks,num_samples, matches("qval$")) %>%
  gather(variable, qval, matches("^D[0-9]"))

peaks_boolean_clean$qval <- as.numeric(peaks_boolean_clean$qval)

```


## Filter peaks that have qval > 0.00001

Use qval >5 as it is -log10

Peak is completely eliminated if it does not pass filter in any sample

```{r}

peaks_filter_gather <- peaks_boolean_clean %>%
  filter(qval > 5)

peaks_filter_wide <- peaks_filter_gather %>%
  select(interval_id,variable,qval) %>%
  spread(variable,qval)

write.table(peaks_filter_wide, file =paste0(workingdir,outdir,"consensus_peakfdr_filtered.csv"), quote = FALSE, row.names = FALSE)

```


```{r}
sessionInfo()
```

