---
title: "Footprinting across p0-M"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Footprinting differences within the shared chromatin landscape of p0-M

```{r message=FALSE}
rm(list=ls())

library(RColorBrewer)
library(tidyverse)
library(genomation)
library(matrixStats)
library(ComplexHeatmap)
library(readxl)
library(broom)
library(patchwork)

```

### Input data

Ouput of bindetect is in `BINDetect_25conditions_arch/bindetect_results.txt`

Motif annotation excel file was downloaded from Vierstra's [motif clustering](https://www.vierstra.org/resources/motif_clustering#downloads)

Archetype to possible genes lookup table `lookup_motif_gene_archetype_21-11-23.csv` is custom made. 

```{r}



Motif_preds_tabl <- read.table("/Users/delasj/Documents/BriscoeLab/project_DV_ATAC_reproduce_analysis/inputs_footprinting/bindetect_results.txt",header=T,sep="\t",stringsAsFactors=F)


```

Clean up table and keep relevant samples.

```{r}

Motif_scores <- Motif_preds_tabl %>%
  select(c("output_prefix" ,"name","motif_id","cluster", "motif_logo","total_tfbs",contains("_score"))) %>%
  select(-contains("_3_mean_score")) %>%
  select(-contains("NMP_mean_score"))


```

### table to group motifs to archetypes
```{r}
archetypes_clusterid <- read_excel("/Users/delasj/Documents/BriscoeLab/project_DV_ATAC_reproduce_analysis/inputs_footprinting/motif_annotations.xlsx",2)  
archtypes_names <- read_excel("/Users/delasj/Documents/BriscoeLab/project_DV_ATAC_reproduce_analysis/inputs_footprinting/motif_annotations.xlsx",1)  

archtypes_2_motifs <- archetypes_clusterid %>% select(c("Cluster_ID","Motif","Database","Consensus")) %>%
  left_join(archtypes_names %>% select(c("Cluster_ID","Name","DBD","Seed_motif")), by="Cluster_ID")
archtypes_2_motifs$Cluster_ID <- as.character(archtypes_2_motifs$Cluster_ID)

```

### Archetype to possible TFs

```{r}
gene_motifs_cluster <- read.csv("/Users/delasj/Documents/BriscoeLab/project_DV_ATAC_reproduce_analysis/inputs_footprinting/lookup_motif_gene_archetype_21-11-23.csv", stringsAsFactors =FALSE,
                                colClasses=c("character","character"))
```

### Gene expression

```{r}

gene_exp <- read.csv("../20-07-30_RNA_DEA/Export_counts/RNA_normCounts_filter1.csv", stringsAsFactors =FALSE )


```



## My colors and other settings

```{r}
# Annotated heatmap with selected colors
hm_colors = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100)

# sample order etc
sorted.DayGate <- c("D3_NMP","D4_1","D4_2","D4_M","D4_3",
                    "D5_1","D5_2","D5_M","D5_3",
                    "D6_1","D6_2","D6_M","D6_3")

sorted.samples <- c("D3_0_NMP","D4_0_1","D4_10_1","D5_0_1","D5_10_1","D6_0_1","D6_10_1",
                   "D4_10_2","D4_100_2","D5_10_2","D5_100_2","D6_10_2","D6_100_2",
                   "D4_100_M","D4_500_M","D5_100_M","D5_500_M","D6_100_M","D6_500_M",
                   "D4_100_3","D4_500_3","D5_100_3","D5_500_3","D6_100_3","D6_500_3")

colorJD <- c("#477d92","#e5a114","#e3602b",
            "#2e525e","#9f7113","#ab4117")

```


## Filter "significant" as the BINDetect volcano plot does

The volcano plots shows
y_min = np.percentile(yvalues[yvalues > 0], 5)	#5% smallest pvalues
x_min, x_max = np.percentile(xvalues, [5, 95])	#5% smallest and largest changes

Filter lowest pval and largest changes PER comparison

```{r sign-motifs}

scores_plotTop <- Motif_preds_tabl %>% 
  select(c("output_prefix",ends_with("_score"))) %>%
  select(-contains("_3_mean_score")) %>%
  select(-contains("NMP_mean_score")) %>%
  select(-starts_with("D4")) %>%
  gather(quant_item, value,ends_with("_score"))

scores_plot_topchanges <- Motif_preds_tabl %>%
  select("output_prefix", ends_with("_change")) %>%
  gather(item_changes, value_change, ends_with("_change")) %>% 
  group_by(item_changes) %>%
  mutate(value_abs_change = abs(value_change)) %>%
  top_frac(0.05,value_abs_change) %>%
  ungroup() %>%
  spread(item_changes, value_change)

scores_plot_topovalues <- Motif_preds_tabl %>%
  select("output_prefix", ends_with("_pvalue")) %>%
  gather(item_pvalue, value_pvalue, ends_with("_pvalue")) %>% 
  group_by(item_pvalue) %>%
  top_frac(-0.05,value_pvalue) %>%
  ungroup() %>%
  spread(item_pvalue, value_pvalue)


```

## Heatmap of top variable or significant

With the top motifs, get heatmap

```{r most-variable-motifs, fig.height=7, fig.width=8}

scores_plot <- Motif_preds_tabl %>% select(c("output_prefix","name","motif_id","cluster", ends_with("_score"))) %>%
  select(-contains("_3_mean_score")) %>%
  select(-contains("NMP_mean_score")) %>%
  select(-starts_with("D4"))

scores_subset <- scores_plot %>% 
  filter(output_prefix %in% as.character(scores_plot_topovalues$output_prefix) | output_prefix %in% as.character(scores_plot_topchanges$output_prefix)) 


## annotate with archetypes again 

## Annotation to color by archetype
scores_cluster_ann <- scores_subset %>%
  left_join(archtypes_2_motifs, by = c("motif_id"="Motif")) %>%
  mutate(Archetype=paste(Cluster_ID,Name,sep = "_")) %>%
  remove_rownames() %>%
  column_to_rownames("output_prefix") %>%
  select("Archetype")

```


## calculate z-scores
```{r zscore-calculate }

scores_subset_plot_z <- scores_subset_hm_z %>%
  as.data.frame() %>%
  rownames_to_column("Archetype")
  
scores_subset_plot_z_gather <- scores_subset_plot_z %>%
  gather("sample","zscore", ends_with("_score")) %>%
  mutate(sample=gsub("_mean_score","", sample),
         sample=factor(sample, levels = sorted.samples)) %>%
  separate(sample,into=c("Day","SAG","Gate"), sep="_", remove=FALSE) %>%
  mutate(Condition=paste(Day,SAG,Gate, sep="_"),
         DaySAG=paste(Day,SAG,sep = "_"),
         DayGate=paste(Day,Gate,sep="_")) %>%
  mutate(DayGate = factor(DayGate, levels = sorted.DayGate))
```



# Correlation with gene expression

Upload the lookup table and the gene expression matrix

```{r gene-exp-tables}

gene_2_clusterID_all <- gene_motifs_cluster %>% select("Cluster_ID","mouse_genename") %>% unique()

gene_2_clusterID <- gene_2_clusterID_all %>% 
  filter(!is.na(mouse_genename))

gene_2_arch <- gene_2_clusterID %>%
  left_join(archtypes_2_motifs, by = "Cluster_ID") %>%
  mutate(Archetype=paste(Cluster_ID,Name,sep = "_"))

```


Clean tables
```{r clean-rna-tables }

gene_exp_preclean <- gene_exp %>%
  rename(genename=X) %>%
  gather(sample, norm_counts, starts_with("D")) %>%
  mutate(sample=gsub("\\.","_", sample)) %>%
  separate(sample,into=c("Day","SAG","Gate","Rep"), sep="_") %>%
  mutate(Condition=paste(Day,SAG,Gate, sep="_")) %>%
  group_by(Condition, genename) %>%
  summarise(ave_count = mean(norm_counts))


# Only keep genes where at least 1 Condition has more than 100 normalized counts (average across reps)
gene_exp_clean <- gene_exp %>%
  rename(genename=X) %>%
  gather(sample, norm_counts, starts_with("D")) %>%
  mutate(sample=gsub("\\.","_", sample)) %>%
  separate(sample,into=c("Day","SAG","Gate","Rep"), sep="_") %>%
  mutate(Condition=paste(Day,SAG,Gate, sep="_")) %>%
  group_by(Condition, genename) %>%
  summarise(ave_count = mean(norm_counts), .groups="keep") %>%
  spread(Condition, ave_count) %>%
  filter_at(vars(starts_with("D")), any_vars(.>100)) %>%
  gather(Condition, ave_count, starts_with("D"))


zscores_arch_clean <- scores_subset_plot_z_gather %>% 
  separate(Archetype, into = c("Cluster_ID","Cluster_name"), sep = "_", remove = FALSE) 

# archtypes that still have genes
gene_2_arch_expressionfiltered = gene_2_arch %>%
  filter(mouse_genename %in% gene_exp_clean$genename)



 
```


To save a long list of correlations

```{r function-all-cor-fits}

get_reg_fit_arch_RNA <- function(x){
  
  clean_x = gsub("/","-",x)
  
  sub_zscores = zscores_arch_clean %>%
  filter(Archetype==x)
  
  sub_genes = gene_2_arch %>% filter(Archetype==x)
  
  sub_exp = gene_exp_clean %>% filter(genename %in% as.character(sub_genes$mouse_genename))
  
  sub_plot = sub_zscores %>%
    left_join(sub_exp, by = c("sample"="Condition")) 

  sub_corr <- sub_plot %>%
    group_by(genename) %>%
    mutate(arch2gene_corr = cor(zscore,ave_count)) %>%
    select(Archetype,Cluster_ID,Cluster_name,genename,arch2gene_corr) %>%
    unique()
  
  sub_fit <- sub_plot %>%
    group_by(genename) %>%
    do(fitgene =  glance(lm(zscore ~ ave_count, data = .))) %>%
    unnest(fitgene) %>%
    mutate(Archetype=x)
  
  sub_fit_corr <- sub_fit %>%
    inner_join(sub_corr, by=c("genename","Archetype"))

  sub_fit_corr
} 

top_arch <- intersect(zscores_arch_clean$Archetype, gene_2_arch_expressionfiltered$Archetype) %>% unique() 

Arch_gene_fit_all <- lapply(top_arch, get_reg_fit_arch_RNA)

Arch_gene_fit_table <- do.call(rbind,Arch_gene_fit_all)

```


### Plot the top Correlated 

Get the top archetypes-gene from Arch_gene_fit_table

Go back to function-all-cor-fits and get the full table


```{r plot-top-corr, fig.width=11}

Top_fit <- Arch_gene_fit_table %>%
  filter(adj.r.squared > 0.6)

  
hist(Arch_gene_fit_table$adj.r.squared)


top_sub_zscores = zscores_arch_clean %>%
  filter(Archetype %in% Top_fit$Archetype)
  
top_sub_genes = gene_2_arch %>% filter(Archetype%in% Top_fit$Archetype & mouse_genename %in% Top_fit$genename) %>%
  select(-Motif,-Database,-Consensus) %>% unique()

top_sub_exp = top_sub_genes %>%
  left_join(gene_exp_clean, by = c("mouse_genename"="genename"))

top_sub_plot = top_sub_zscores %>%
  left_join(top_sub_exp, by = c("Archetype","Condition"))


ggplot(top_sub_plot, aes(x=ave_count, y=zscore, group=mouse_genename)) +
  geom_smooth(method=lm, formula = y ~ x) +
  geom_point(aes(color=DayGate)) +
  scale_color_manual(values = colorJD) +
  facet_wrap(Archetype ~ mouse_genename, scales = "free_x", ncol = 6) +
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1)


```


### Plot individual ones 


```{r}
archetype = "5_HD/5"


test_whichzscores = scores_subset_plot_z_gather %>% separate(Archetype, into = c("Cluster_ID","Cluster_name"), sep = "_", remove = FALSE) %>%
  filter(Archetype==archetype)
  

test_whichgenes <- gene_2_arch %>% filter(Archetype==archetype)

test_whichgene_exp <- gene_exp_clean %>% filter(genename %in% as.character(test_whichgenes$mouse_genename))

test_corr <- test_whichzscores %>%
  left_join(test_whichgene_exp, by = c("sample"="Condition")) %>%
  group_by(genename) %>%
  mutate(arch2gene_corr = cor(zscore,ave_count))

ggplot(test_corr, aes(x=ave_count, y=zscore)) +
  geom_smooth(method=lm) +
  geom_point(aes(color=DayGate, shape=SAG)) +
  scale_color_manual(values = colorJD) +
  facet_wrap( ~ genename, scales = "free_x")+  
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1)


```


```{r}
archetype = "281_PAX/1"


test_whichzscores = scores_subset_plot_z_gather %>% separate(Archetype, into = c("Cluster_ID","Cluster_name"), sep = "_", remove = FALSE) %>%
  filter(Archetype==archetype)
  

test_whichgenes <- gene_2_arch %>% filter(Archetype==archetype)

test_whichgene_exp <- gene_exp_clean %>% filter(genename %in% as.character(test_whichgenes$mouse_genename))

test_corr <- test_whichzscores %>%
  left_join(test_whichgene_exp, by = c("sample"="Condition")) %>%
  group_by(genename) %>%
  mutate(arch2gene_corr = cor(zscore,ave_count))

ggplot(test_corr, aes(x=ave_count, y=zscore)) +
  geom_smooth(method=lm) +
  geom_point(aes(color=DayGate, shape=SAG)) +
  scale_color_manual(values = colorJD) +
  facet_wrap( ~ genename, scales = "free_x")+  
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1)


```

```{r}
archetype = "63_Ebox/CAGATGG"


test_whichzscores = scores_subset_plot_z_gather %>% separate(Archetype, into = c("Cluster_ID","Cluster_name"), sep = "_", remove = FALSE) %>%
  filter(Archetype==archetype)
  

test_whichgenes <- gene_2_arch %>% filter(Archetype==archetype)

test_whichgene_exp <- gene_exp_clean %>% filter(genename %in% as.character(test_whichgenes$mouse_genename))

test_corr <- test_whichzscores %>%
  left_join(test_whichgene_exp, by = c("sample"="Condition")) %>%
  group_by(genename) %>%
  mutate(arch2gene_corr = cor(zscore,ave_count))

ggplot(test_corr, aes(x=ave_count, y=zscore)) +
  geom_smooth(method=lm) +
  geom_point(aes(color=DayGate, shape=SAG)) +
  scale_color_manual(values = colorJD) +
  facet_wrap( ~ genename, scales = "free_x")+  
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1)


```

## Plot the RNA and the motif scores side by side for fig

```{r, fig-RNA-motifs}
#ordered by gate
sorted.DayGate2 <- c("D3_NMP","D4_1","D5_1","D6_1",
                     "D4_2","D5_2","D6_2",
                     "D4_M","D5_M","D6_M",
                     "D4_3","D5_3","D6_3")
#ordered by gate
colorJD2 <- c("#6da4ba","#477d92","#2e525e",
              "#f0be56","#e5a114","#9f7113",
              "#ec936f","#e3602b","#ab4117",
              "#5bb357","#009640","#044a23")
# 3 colors ordered by gate
colorJD3 <- c("#477d92","#e5a114","#e3602b","#009640")

# scores_for_plot <- scores_subset_cluster %>% select(c("Archetype", ends_with("score"))) %>%
#   gather("sample","score", ends_with("_score")) %>%
#   mutate(sample=gsub("_mean_score","", sample),
#          sample=factor(sample, levels = sorted.samples)) %>%
#   separate(sample,into=c("Day","SAG","Gate"), sep="_", remove=FALSE) %>%
#   mutate(Condition=paste(Day,SAG,Gate, sep="_"),
#          DaySAG=paste(Day,SAG,sep = "_"),
#          DayGate=paste(Day,Gate,sep="_")) %>%
#   mutate(DayGate = factor(DayGate, levels = sorted.DayGate2))


# I need all the scores, not just D5 D6
scores_subset_alldays <- Motif_preds_tabl %>%  
  select(c("output_prefix","name","motif_id","cluster", ends_with("_score"))) %>% 
  select(-contains("_3_mean_score")) %>%
  select(-contains("NMP_mean_score"))
  #select(-starts_with("D4")) 
  

scores_subset_cluster_top_alldays = scores_subset_alldays %>%
  left_join(archtypes_2_motifs, by = c("motif_id"="Motif")) %>%
  mutate(Archetype=paste(Cluster_ID,Name,sep = "_")) %>%
  gather(sample, mean_score, ends_with("_score")) %>%
  group_by(output_prefix) %>%
  mutate(score_var = var(mean_score)) %>%
  ungroup() %>% group_by(Archetype) %>%
  top_n(1,score_var) %>%
  select(-score_var, -output_prefix, -name, -motif_id, -cluster, -Database) %>% # eliminate the name of the motif
  unique() %>%
  mutate(sample=gsub("_mean_score","", sample),
         sample=factor(sample, levels = sorted.samples)) %>%
  separate(sample,into=c("Day","SAG","Gate"), sep="_", remove=FALSE) %>%
  mutate(Condition=paste(Day,SAG,Gate, sep="_"),
         DaySAG=paste(Day,SAG,sep = "_"),
         DayGate=paste(Day,Gate,sep="_")) %>%
  mutate(DayGate = factor(DayGate, levels = sorted.DayGate2))


archetype = "2_HD/2"
gene="Nkx6-1"

sub_zscores = scores_subset_cluster_top_alldays %>% separate(Archetype, into = c("Cluster_ID","Cluster_name"), sep = "_", remove = FALSE) %>%
  filter(Archetype==archetype)

sub_counts <- gene_exp %>% filter(X == gene)


sub_counts_clean <- sub_counts %>%
  select(-starts_with("D3")) %>%
  rename(genename=X) %>%
  gather(sample, norm_counts, starts_with("D")) %>%
  mutate(sample=gsub("\\.","_", sample)) %>%
  separate(sample,into=c("Day","SAG","Gate","Rep"), sep="_") %>%
  filter(Gate != "3") %>%
  mutate(Condition=paste(Day,SAG,Gate, sep="_"),
         DayGate=paste(Day,Gate,sep="_")) %>%
  mutate(DayGate = factor(DayGate, levels = sorted.DayGate2),
         Condition=factor(Condition, levels = sorted.samples))


p1 <- ggplot(sub_counts_clean, aes(x=DayGate, y=norm_counts, group=Gate)) +
  geom_smooth(se = FALSE, method = lm, color="grey30") +
  geom_point(aes(fill=DayGate), shape=21, size=2) +
  scale_fill_manual(values = colorJD2) +
  facet_wrap( ~ genename, scales = "free_x") +
  ylab("Normalized RNA counts") +
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1,axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p2 <- ggplot(sub_zscores, aes(x=DayGate, y=mean_score, group=Gate)) +
  geom_smooth(se = FALSE, method = lm, color="grey30") +
  geom_point(aes(fill=DayGate), shape=21, size=2) +
  scale_fill_manual(values = colorJD2) +
  facet_wrap( ~ Cluster_name, scales = "free_x") +
  ylab("Motif score") +
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1,axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p2 + p1


```



```{r, fig-RNA-motifs2}
archetype = "63_Ebox/CAGATGG"
gene="Olig2"

sub_zscores = scores_subset_cluster_top_alldays %>% separate(Archetype, into = c("Cluster_ID","Cluster_name"), sep = "_", remove = FALSE) %>%
  filter(Archetype==archetype)

sub_counts <- gene_exp %>% filter(X == gene)


sub_counts_clean <- sub_counts %>%
  select(-starts_with("D3")) %>%
  rename(genename=X) %>%
  gather(sample, norm_counts, starts_with("D")) %>%
  mutate(sample=gsub("\\.","_", sample)) %>%
  separate(sample,into=c("Day","SAG","Gate","Rep"), sep="_") %>%
  filter(Gate != "3") %>%
  mutate(Condition=paste(Day,SAG,Gate, sep="_"),
         DayGate=paste(Day,Gate,sep="_")) %>%
  mutate(DayGate = factor(DayGate, levels = sorted.DayGate2),
         Condition=factor(Condition, levels = sorted.samples))


p1 <- ggplot(sub_counts_clean, aes(x=DayGate, y=norm_counts, group=Gate)) +
  geom_smooth(se = FALSE, method = lm, color="grey30") +
  geom_point(aes(fill=DayGate), shape=21, size=2) +
  scale_fill_manual(values = colorJD2) +
  facet_wrap( ~ genename, scales = "free_x") +
  ylab("Normalized RNA counts") +
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1,axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p2 <- ggplot(sub_zscores, aes(x=DayGate, y=mean_score, group=Gate)) +
  geom_smooth(se = FALSE, method = lm, color="grey30") +
  geom_point(aes(fill=DayGate), shape=21, size=2) +
  scale_fill_manual(values = colorJD2) +
  facet_wrap( ~ Cluster_name, scales = "free_x") +
  ylab("Motif score") +
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1,axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


gene="Neurog2"
sub_counts <- gene_exp %>% filter(X == gene)


sub_counts_clean <- sub_counts %>%
  select(-starts_with("D3")) %>%
  rename(genename=X) %>%
  gather(sample, norm_counts, starts_with("D")) %>%
  mutate(sample=gsub("\\.","_", sample)) %>%
  separate(sample,into=c("Day","SAG","Gate","Rep"), sep="_") %>%
  filter(Gate != "3") %>%
  mutate(Condition=paste(Day,SAG,Gate, sep="_"),
         DayGate=paste(Day,Gate,sep="_")) %>%
  mutate(DayGate = factor(DayGate, levels = sorted.DayGate2),
         Condition=factor(Condition, levels = sorted.samples))

p3 <- ggplot(sub_counts_clean, aes(x=DayGate, y=norm_counts, group=Gate)) +
  geom_smooth(se = FALSE, method = lm, color="grey30") +
  geom_point(aes(fill=DayGate), shape=21, size=2) +
  scale_fill_manual(values = colorJD2) +
  facet_wrap( ~ genename, scales = "free_x") +
  ylab("Normalized RNA counts") +
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1,axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


gene="Neurod1"
sub_counts <- gene_exp %>% filter(X == gene)


sub_counts_clean <- sub_counts %>%
  select(-starts_with("D3")) %>%
  rename(genename=X) %>%
  gather(sample, norm_counts, starts_with("D")) %>%
  mutate(sample=gsub("\\.","_", sample)) %>%
  separate(sample,into=c("Day","SAG","Gate","Rep"), sep="_") %>%
  filter(Gate != "3") %>%
  mutate(Condition=paste(Day,SAG,Gate, sep="_"),
         DayGate=paste(Day,Gate,sep="_")) %>%
  mutate(DayGate = factor(DayGate, levels = sorted.DayGate2),
         Condition=factor(Condition, levels = sorted.samples))

p4 <- ggplot(sub_counts_clean, aes(x=DayGate, y=norm_counts, group=Gate)) +
  geom_smooth(se = FALSE, method = lm, color="grey30") +
  geom_point(aes(fill=DayGate), shape=21, size=2) +
  scale_fill_manual(values = colorJD2) +
  facet_wrap( ~ genename, scales = "free_x") +
  ylab("Normalized RNA counts") +
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1,axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



p2+p1+p3+p4


```


```{r, fig-RNA-motifs3}
archetype = "281_PAX/1"
gene="Pax6"

sub_zscores = scores_subset_cluster_top_alldays %>% separate(Archetype, into = c("Cluster_ID","Cluster_name"), sep = "_", remove = FALSE) %>%
  filter(Archetype==archetype)

sub_counts <- gene_exp %>% filter(X == gene)


sub_counts_clean <- sub_counts %>%
  select(-starts_with("D3")) %>%
  rename(genename=X) %>%
  gather(sample, norm_counts, starts_with("D")) %>%
  mutate(sample=gsub("\\.","_", sample)) %>%
  separate(sample,into=c("Day","SAG","Gate","Rep"), sep="_") %>%
  filter(Gate != "3") %>%
  mutate(Condition=paste(Day,SAG,Gate, sep="_"),
         DayGate=paste(Day,Gate,sep="_")) %>%
  mutate(DayGate = factor(DayGate, levels = sorted.DayGate2),
         Condition=factor(Condition, levels = sorted.samples))


p1 <- ggplot(sub_counts_clean, aes(x=DayGate, y=norm_counts, group=Gate)) +
  geom_smooth(se = FALSE, method = lm, color="grey30") +
  geom_point(aes(fill=DayGate), shape=21, size=2) +
  scale_fill_manual(values = colorJD2) +
  facet_wrap( ~ genename, scales = "free_x") +
  ylab("Normalized RNA counts") +
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1,axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p2 <- ggplot(sub_zscores, aes(x=DayGate, y=mean_score, group=Gate)) +
  geom_smooth(se = FALSE, method = lm, color="grey30") +
  geom_point(aes(fill=DayGate), shape=21, size=2) +
  scale_fill_manual(values = colorJD2) +
  facet_wrap( ~ Cluster_name, scales = "free_x") +
  ylab("Motif score") +
  theme_bw(base_size = 12) +
  theme(aspect.ratio = 1,axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p2+p1
```



```{r}
sessionInfo()
```
