---
title: "Overlap of ATAC clusters with regions opened in Foxa2 inducible or control"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Analysis

Overlap between intervals of different clusters and the up or down significant intervals in the inducible Foxa2

```{r message=FALSE}
# Load all packages
rm(list=ls())

library(tidyverse)
library(RColorBrewer)

```

### Set dirs
```{r}

workingdir="/Users/delasj/Documents/BriscoeLab/project_DV_ATAC_reproduce_analysis/"
subworkinput="inputs_ifoxa2-atac_2_overlaps/"

```


### Define colors 

```{r}

cluster_colors <- c("#878787","#ff284d","#ff93a6","#96c283","#55863f","#edb273","#e69138","#b8742c","#3d85c6")
  
```


## Intersects of the clusters and elements opened by Foxa2 or control

```{r fig.width = 13, fig.height=12}

beds_import <- lapply(list.files(path = paste0(workingdir,subworkinput), pattern="*.bed", full.names=TRUE),function(x){
    data <- read.table(x, header = FALSE)
    data$ChIP <- gsub("__within__.*","",x)
    data$ChIP <- gsub(".*/","",data$ChIP)
    data$cluster <- gsub(".*__within__","",x)
    data$cluster <- gsub("\\.bed","",data$cluster)
    data
})

all_intersects <- do.call(rbind,beds_import)
```

## Import the overal beds
### Full beds for the clusters 

```{r}

fullbeds_import <- lapply(list.files(path = paste0(workingdir,"outputs_cats-atac_5_kmeans"), pattern="*bed", full.names=TRUE),function(x){
    data <- read.table(x, header = FALSE)
    data$cluster <- gsub("Intervals_","",x)
    data$cluster <- gsub("\\.bed","",data$cluster)
    data$cluster <- gsub(".*/","",data$cluster)
    data
})

full_beds <- do.call(rbind,fullbeds_import)


```




### Full beds for Foxa2 diff acc
```{r}

fullchip_import <- lapply(list.files(path = paste0(workingdir,"outputs_ifoxa2-atac_1"), pattern="*.bed*", full.names=TRUE),function(x){
    data <- read.table(x, header = FALSE)
    data <- data %>% select(c("V1","V2","V3","V4"))
    data$ChIP <- gsub("\\.bed","",x)
    # data$ChIP <- gsub("\\.consensus_peaks\\.bed","",data$ChIP)
    data$ChIP <- gsub(".*/","",data$ChIP)
    data
})

full_chip <- do.call(rbind,fullchip_import)


```


## Fraction of Cluster occupied by ChIP
## Fraction of Chip calls within each cluster

Bedtools intersect -wa will report a (the interval in cluster) multiple times if there are multiple overlaps. 

```{r}
intersects_sums = all_intersects %>%
  select(V4,ChIP,cluster) %>% 
  unique() %>%
  group_by(cluster, ChIP) %>%
  summarise(count=n(), .groups = 'keep')

cluster_sums = full_beds %>%
  group_by(cluster) %>%
  summarise(total=n(), .groups = 'keep')

chip_sums = full_chip %>%
  group_by(ChIP) %>%
  summarise(total=n(), .groups = 'keep')

intersect_full_counts <- intersects_sums %>%
  left_join(cluster_sums, by = "cluster") %>%
  dplyr::rename(cluster_n=total) %>%
  left_join(chip_sums, by = "ChIP") %>%
  dplyr::rename(chip_n=total)


intersect_full_prop <- intersect_full_counts %>%
  mutate(fraction_of_cluster=count/cluster_n,
         fraction_of_chip=count/chip_n,
         cluster_ChIP=paste(cluster, ChIP, sep = "__"))
  

```


## Proportions bar plots
### plot for figure

Fraction of elements in the cluster that overlap with elements in Foxa2 Up or Down

```{r}

ggplot(intersect_full_prop %>% 
         mutate(cluster=factor(cluster, levels = c("Cluster_1","Cluster_3","Cluster_2","Cluster_9",
                                                   "Cluster_6","Cluster_7","Cluster_8","Cluster_4","Cluster_5"))), 
       aes(x=cluster, y=fraction_of_cluster, fill=cluster)) +
  geom_bar(stat = "identity", position = "stack") + 
  facet_wrap( ~ ChIP, scales = "free_y") +
  theme_bw() +
  scale_fill_manual(values = cluster_colors[c(1,3,2,9,6,7,8,4,5)]) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```







```{r}
sessionInfo()
```

